Large-scale harmonic patterns clustering

### 2010-03-11 Paper/Paper.tex
improvements

Other large-scale work use summarization of the features. We mention 
gaussian mixtures \cite{Mandel2005}, aggregate features \cite{Bergstra2006a}
and linear dynamical systems \cite{Barrington2009a}. The chroma representation
is close from an harmonic point of view to the actual music, in particular
we can ``listen'' to the patches (see Section \ref{sec:experiments}).
This is an advantage for analyzing a large data set, and in accordance with our goal
we restrain ourselves from using generative models to compress our features.

@@

We try to normalize the patterns regarding the key. We roll the matrix
so the row with the more energy is always the lowest one. This can easily be
seen in Figure \ref{fig:code}. We considered the same technique for rolling
along the other axis and be invariant regarding the downbeat. However, it
seems less robust to noise. Note that we roll each pattern independentely.
Also, we imply that the pitch envelope is important rather than the actual 
pitches.

\subsection{Codebook Learning}
We use an online version of the vector quantization algorithm 
\cite{Gersho1991}. This can also be seen as an online Kmeans.
The idea is as follow: take a sample from your data, find the closest
code in the codebook, bring that code closer to the sample by some amount.
The details are explained as Algorithm \ref{algo:vq}. note that this
algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
Thus it scales linearly to any size of data set. The setting of the learning
rate might become trickier, but this is not fundamentally a problem.

@@

We use the codebook to encode our testing set (see Section \ref{sec:data}).
Each pattern is encoded with only one code. We can measure the average
distance between a pattern and its encoding. We can also measure the use
of the different codes.

@@

\subsection{Data Size}
We are interesting to see how the amount of training data changes
the quality of our codebook.
We experiment with different data sizes for a fixed size of codebook and a 
fixed code length. The data points are chose randomly from the whole
data set.
In the experiment a pattern size correspond to $1$ bar resized to $4$
beats, thus a pattern size of $12$x$4$. The codebook size is $100$.
We measure the distortion on our test set. We hope for some convergence.
See results in Figure \ref{fig:sizes}.

@@

\subsection{Codebook Size}
We experiment with the size of the codebook and its affect on the
performance of the encoding. Results are shown in Table \ref{tab:cbsize}.
Note that we trained on a number of samples depending on the codebook
size, i.e. we used $500$ samples per code to learn. This is consistent
with our findings in Figure \ref{fig:sizes}.

@@

\subsection{Visualization}
Interpreting a codebook of size larger than $100$ is not obvious.
We can use an algorithm like locally linear embedding \cite{Roweis2000} (LLE) 
or multidimensional scaling \cite{Kruskal1964} to visualize the data in
2 dimensions. We use the former\footnote{implementation: 
\url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}, the
result can be seen in Figure \ref{fig:lle}.

@@

\section{Links with NMF}\label{sec:nmf}
In this section we discuss links with non-negative matrix factorization
\cite{Lee2000} (NMF), another method to build a codebook representing some data.
NMF finds two matrices $H$ and $W$ that approximate some data $D$ (one
sample per line) by
\[
D = W \times H + \epsilon
\]
where $\epsilon$ is some noise and $W$ and $H$ only have non-negative 
elements. If $W$ is $\#samples \times n$, $n$ is the codebook size.

@@

>>>>

Other large-scale work use summarization of the features. We mention 
gaussian mixtures \cite{Mandel2005}, aggregate features \cite{Bergstra2006a}
and linear dynamical systems \cite{Barrington2009a}. The chroma representation
is close from an harmonic point of view to the actual music, in particular
we can ``listen'' to the patches (see Section \ref{sec:experiments}).
This is an advantage for analyzing a large data set, thus
we restrain ourselves from using generative models to compress our features.

@@

We try to normalize the patterns in respect to the key. We roll the matrix
so the row with the more energy is always the lowest one. This can easily be
seen in Figure \ref{fig:code}. We considered the same technique for rolling
along the other axis and be invariant regarding the downbeat. However, it
seems less robust to noise. Note that we roll each pattern independentely.
Also, this implies that the pitch envelope is important rather than the actual 
pitches.

\subsection{Codebook Learning}
We use an online version of the vector quantization algorithm 
\cite{Gersho1991}. This can also be seen as an online K-means.
The idea is as follow: take a sample from your data, find the closest
code in the codebook, bring that code closer to the sample by some amount.
The details are explained in Algorithm \ref{algo:vq}. Note that this
algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
Thus it scales linearly to any size of data set. The setting of the learning
rate might become trickier, but this is not fundamentally a problem.

@@

We use the codebook to encode our testing set (see Section \ref{sec:data}).
Each pattern is encoded with only one code. We can measure the average
distance between a pattern and its encoding. We can also measure the use
of the different codes, i.e. the size of the cluster around a code.

@@

\subsection{Data Size}
We are interesting to see how the amount of training data changes
the quality of our codebook.
We experiment with different data sizes for a fixed size of codebook and a 
fixed code length. The data points are chose randomly from the whole
data set.
In the experiment a pattern size correspond to $1$ bar resized to $4$
beats, thus a pattern size of $12$x$4$. The codebook size is $100$.
We measure the distortion on our test set. 
See results in Figure \ref{fig:sizes}.

@@

\subsection{Codebook Size}
We experiment with the size of the codebook and its affect on the
performance of the encoding. Results are shown in Table \ref{tab:cbsize}.
Note that we trained on a number of samples depending on the codebook
size, i.e. we used $500$ samples per code to learn. This is consistent
with our findings from Figure \ref{fig:sizes}.

@@

\subsection{Visualization}
Interpreting a codebook of size larger than $100$ is not obvious.
We can use an algorithm like locally linear embedding \cite{Roweis2000} (LLE) 
or multidimensional scaling \cite{Kruskal1964} to visualize the data in
2 dimensions. Here we use the former\footnote{implementation: 
\url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}, the
result can be seen in Figure \ref{fig:lle}.

@@

\section{Links with NMF}\label{sec:nmf}
In this section we discuss links with non-negative matrix factorization
\cite{Lee2000} (NMF), another method to build a codebook representing some data.
NMF finds two matrices $H$ and $W$ that approximate some data $D$ (one
sample per line) by
\begin{eqnarray}
D \approx W^{m \times k} H^{k \times n}
\end{eqnarray}
where $\epsilon$ is some noise and $W$ and $H$ only have non-negative 
elements. If $W$ is $\#samples \times k$, $k$ is the codebook size.

@@

Moving on to larger scales, we could imagine merging codes together and
splitting some on two if the clusters become too small or too large.We could
also cluster the codebook itself, in a similar fashion as hierarchical
gaussian mixtures \cite{Vasconcelos2001}.
